---
title: "Disinformation analysis YouTube"
author: "Maggie Mazza, Casper Polhaar"
date: "2026-05-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
#| label: load packages
#| warning: false
#| message: false

#load neccessary packages
library(tidyverse)
library(tinytex)
library(tidytext)
library(knitr)
library(corrr)

# Read the datasets
# (Ensure these files are in your working directory)
video_meta <- read_csv("youtube_haitian_disinformation_video_meta.csv")
comment_meta <- read_csv("youtube_haitian_disinformation_comment_reply_metadata.csv")

# View the main dataset
glimpse(video_meta)
```

## Data preparation and comment counting

```{r}
#| label: data-cleaning

# Count comments per video
comment_counts <- comment_meta %>%
  group_by(video_id) %>% 
  summarise(comment_count = n(), .groups = 'drop')

# Merge, clean, and prep the dataset
video_prep <- video_meta %>%
  # Merge the comment counts
  left_join(comment_counts, by = "video_id") %>%
  # If a video had no comments in the comment file, replace NA with 0
  mutate(comment_count = replace_na(comment_count, 0)) %>%
  # Remove irrelevant variables
  select(-any_of(c("date", "tags", "comments", "transcript", "transcript_str"))) %>%
  # Create the dichotomous variable (filled with NA for now so you can code it)
  mutate(video_debunking = NA_real_)

# View the prepped data
head(video_prep)
```

## Coding the video_debunking variable

Since there were only 23 videos, manual coding was a very feasible option, and therefore chosen. This ensures no errors as a result of automation in the variable which is to be created below. The corresponding video_id's where found by opening dataframe 'video_prep' in the environment.

|  |  |
|-------------------------------------------|-----------------------------|
| Title | Debunking / Not debunking |
| Why Trump keeps pushing debunked claims about Haitian immigrants | Debunking |
| Springfield mayor responds to JD Vance’s claims about Haitian immigrants | Not debunking |
| Trump, Harris face off at the ABC News Presidential Debate  | Not debunking |
| Trump amplifies false claims about Haitian immigrants eating pets  | Debunking |
| Springfield, Ohio, man reported Haitian migrants snatching geese out of park just 2 weeks ago  | Not debunking |
| Springfield resident describes impact of Haitian migrants on community  | Not debunking |
| Springfield neighbors describe immigrant situation; say they believe pet-snatching story  | Not debunking |
| JD Vance exposed for knowingly lying about ‘pet-eating’ rumors in Springfield, OH  | Debunking |
| ‘You think this is the last cup of B.S.?’: Steele slams Trump campaign for lying about migrants  | Debunking |
| Ohio residents rage at officials over Haitian migrant surge: ‘No vacancy’  | Not debunking |
| Trump spreads more lies about Springfield, Ohio  | Debunking |
| Springfield, Ohio resident: This is a dystopian nightmare  | Not debunking |
| Haitian community in Springfield, Ohio, under threat after debunked comments  | Debunking |
| Reporter Presses JD Vance About ‘Debunked False Claims’ About Haitians In Springfield, Ohio  | Debunking |
| Criminal charges filed by Haitian nonprofit against Trump and JD Vance  | Debunking |
| ‘NO DOGS OR CATS’ Haitian immigrant pleads with Trump and Springfield, Ohio residents  | Debunking |
| Missing cat at center of Springfield conspiracy theory found hiding in basement  | Debunking |
| Trump falsely claims Haitian migrants are eating pets in Ohio  | Debunking |
| JD Vance claims on Ohio’s Haitian immigrants eating pets debunked  | Debunking |
| Dana Bash and JD Vance clash over baseless claims about Haitian immigrants (Part 1/2)  | Debunking |
| JD Vance reacts to Springfield father saying he politicized his son’s death (Part 2/2)  | Not debunking |
| JD Vance defends amplifying false claims about immigrants  | Debunking |
| Residents in Springfield, Ohio react to accusations of immigrants abducting and eating pets  | Not debunking |

```{r}
#| label: manual-coding-and-factors

# 1. Manually list the exact video IDs that are DEBUNKING
# (Replace these placeholders with your actual video IDs)
debunking_videos <- c("xc0IkVmiN8o", "flRF5OQrURI", "6W0mzgcDNvY", "eytlX2OTKPs", "Ru34gTyHKlw", "0VSV99PXbbw", "ra2PLZlP1o0", "AuItvjZ1Tro", "oBCSCrhnfu0", "mr5Bv-evY9A", "DPCXxNXraKw", "M547oAXe3O4", "i0IvIekwz3c", "LKauI2Gtj4g")

# 2. Apply the manual coding and create the factor
video_final <- video_prep %>%
  mutate(
    # First, create the numeric version for your correlation analysis
    video_debunking_num = ifelse(video_id %in% debunking_videos, 0, 1),
    
    # Next, create the factor version for your tables and summaries
    video_debunking_factor = factor(video_debunking_num,
                                    levels = c(0, 1),
                                    labels = c("debunking", "not debunking"))
  )

# View a quick summary to make sure your factor applied correctly
summary(video_final$video_debunking_factor)
```

## Summarising engagement

```{r}
#| label: summary-table

# Use the factor version so the table reads "debunking" / "not debunking"
overview_table <- video_final %>%
  select(title, video_debunking_factor, views, likes, comment_count)

kable(overview_table, 
      caption = "Overview of Video Engagement",
      format.args = list(big.mark = ","))
```

## Analysing the correlation

```{r}
#| label: correlation-analysis

# Use the numeric version so the correlation math works
correlation_data <- video_final %>%
  select(video_debunking_num, views, likes, comment_count) %>%
  drop_na()

correlation_results <- correlation_data %>%
  correlate(method = "spearman", diagonal = 1) 

kable(correlation_results, 
      digits = 3, 
      caption = "Spearman Correlation Matrix")
```
